---
title: "Power analysis for VR study"
author: "Jarad Niemi"
date: "March 9, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The goal of this project to assess decision making by police officers under
stress. 
Specifically, we would like to understand the relationship between individual
police office personality, situation contextual information, and situation 
stress with decision making timing and accuracy. 
Toward this end, we have developed 3 virtual reality (VR) scenarios:

1. possible gun shot investigation
1. something else
1. office being targeted 

These VR scenarios are designed to increase in stress levels. 
Each VR scenario will have 2 alternative pieces of contextual information,
e.g. the SEC of the neighborhood.
Under each scenario, we will record a variety of responses including time until
a decision is made, decision choice from a set of options, and cortisol level
in the police officer. In addition, the police officers will be given a questionairre to address their
personality traits.

The primary purpose of this document is to explore and suggest a design for the 
VR study and to assess power for the study. 

## Design

The experiment will be conducted at 5 locations across the US. 
Due to the similarity in scenarios each office will never see a VR scenario 
that only differs by contextual information. 
There are two designs main designs that make sense to me:

- Randomized block design
- Randomized block cross-over design
 
We may additionally want to block police officers by personality traits in 
order to reduce the confounding of personality traits with the VR scenarios 
the officers experience.
 
### Randomized block design

In this design, the location would be a block and each police office would see
1 VR scenario. 
Thus, there would be a total of 6 distinct VR scenarios (3 scenarios each with
2 different contextual versions). 
Thus, we should think about putting officers in multiples of 6 at each location.
We should probably also block by time of day, so that we run 6 different 
scenarios for the first 6 officers and then repeat that process throughout the
day. 
Perhaps we could randomize the 6 scenarios according to a latin square in order
to reduce the effect of ordering (I'm not sure this is really necessary). 

If we additionally want to block on officer personality, 
we should group officers by personality into X groups and then run the different
personalities in the same scenario at the same time. 
This has the downside of confounding scenario with time of day. 


### Randomized block cross-over design

If we want to increase the number of VR scenarios each officer sees (due to 
training or learning by the officer or to more accurately measure stress level),
then we can consider a cross-over design where each office sees 3 VR scenarios:
on each of the scenarios listed above with a certain set of contextual 
information.
Considering all possible permuations (scenario, contextual information, and 
order) an officer could see, there are 48 (=6x4x2) different options. 
Consider blocking by 48 individuals: one for each of these scenarios. 

If we want to incorporate blocking by personality traits, 
then these 48 individuals should have similar personality traits. 
This seems hard to enforce. 

```{r, echo=FALSE, eval=FALSE}
library(dplyr)

period1_treatments = c("A1","A2")
period2_treatments = c("B1","B2")
period3_treatments = c("C1","C2")
treatments = c(period1_treatments, 
         period2_treatments,
         period3_treatments)

d = expand.grid(p3 = treatments,
                p2 = treatments,
                p1 = treatments) %>%
  select(p1,p2,p3) %>%
  filter( substr(p1,1,1) != substr(p2,1,1) & 
            substr(p1,1,1) != substr(p3,1,1) & 
            substr(p2,1,1) != substr(p3,1,1) )
```


## Analysis

There will be V VR units utilized in the experiment. 
We should consider the effect of the VR unit.



## Additional questions

- How many VR units will there be?
- How many VR scenarios should an officer see? 

    - Use 1 if we want to maximize the amount of personality trait information we have
    - Use 3 if there is significant training or learning associated with the VR 
  
- What is the most important question?

    - trait affect on decision making (use RCBD)
    - contextual information on decision making (use RCBD. something else?)
    - stress level on decision making (use Cross-over)
  
- How many officers can we reasonably put through the VR scenarios?

    - Does this depend on how many scenarios an officer sees?
  
- How can we categorize officers by personality questionairre?

    - How many groups will there be?
    
- How much carryover do we expect for a police officer in a subsequent 
  